---
title: "index"
output: html_document
---

```{r setup, warning=FALSE, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(leaflet)
library(tigris)

data <- read.csv('../data/global_dataset.csv') 
gdp_data <- read.csv('../data/gdp_dataset.csv') 
raw_state_data <- read.csv('../data/us_state_vaccinations.csv') 

# filtered data which have complete disaggregated data
data <- data %>%
  filter(
    Case...death.data.by.sex. == "Yes"
  )

#select 2020 GDP values
gdp_data <- gdp_data %>%
  select(Country.Name, X2020)

# print column names
data_colnames <- data %>%
  colnames()

# remove NA
data <- na.omit(data)

# change proportion of deaths from string to int

# remove % sign
data$Deaths....male. <- gsub("%$","",data$Deaths....male.)
data$Deaths....female. <- gsub("%$","",data$Deaths....female.)

data$Cases....male. <- gsub("%$","",data$Cases....male.)
data$Cases....female. <- gsub("%$","",data$Cases....female.)
# str to int
data$Deaths....male. <- strtoi(data$Deaths....male.)
data$Deaths....female. <- strtoi(data$Deaths....female.)

data$Cases....male. <- strtoi(data$Cases....male.)
data$Cases....female. <- strtoi(data$Cases....female.)

#merge data
data <- data %>%
  left_join(gdp_data, by = c("Country" = "Country.Name"))

# mean proportion of deaths / cases by gender
female_prop_deaths_mean <- mean(data$Deaths....female.)
male_prop_deaths_mean <- mean(data$Deaths....male.)
female_prop_cases_mean <- mean(data$Cases....female.)
male_prop_cases_mean <- mean(data$Cases....male.)
```
We found that the mean for female deaths is `r `female_prop_deaths_mean`and for male deaths is `r male_prop_deaths_mean`. The mean for female cases is `r female_prop_cases_mean` and for male cases is `r male_prop_cases_mean`.

```{r Distribution of Variables1, warning=FALSE, echo=FALSE} 
# plot for distribution of deaths and case proportions for Females
Gender_deaths_plot <- ggplot(data = data, aes(x = Cases....female.,
                                              y = Deaths....female.)) +
  geom_point() +
  labs(title = "Proportion of Deaths vs Proportion of Cases for Females", x = "Proportion of confirmed cases Female (%)", y = "Proportion of deaths in confirmed cases Female (%)")
Gender_deaths_plot
``` 

This chart shows the proportion of deaths and cases for females. This graph shows that while it is closer to even (50%) some countries have proportions that are significantly lower for females. This may be disrupted by testing access for females in those countries. 

```{r Distribution of Variables2, warning=FALSE, echo=FALSE} 
# plot for distribution of deaths and case proportions for Females
Male_deaths_plot <- ggplot(data = data, aes(x = Cases....male.,
                                              y = Deaths....male.)) +
  geom_point() +
  labs(title = "Proportion of Deaths for Males vs Proportion of Cases for Males", x = "Proportion of confirmed cases Male (%)", y = "Proportion of deaths in confirmed cases Male (%)")
Male_deaths_plot
``` 
We created this chart in order to be able to compare it to the "Proportion of Deaths for Females vs Proportion of Cases for Females" chart. As you can see, the proportion of confirmed cases for males goes above 60%, while for women, it plateaus at 60%. Due to the higher confirmed cases, it makes sense that the proportion of deaths in men will be greater than women. This also means that more men are coming into contact with COVID-19, whether it be from work or even not wearing a mask.

```{r Relationships between Variables1, warning=FALSE, echo=FALSE} 
top_15_country_deaths <- data %>%
  arrange(desc(Deaths.where.sex.disaggregated.data.is.available)) %>%
  select(Country, Deaths.where.sex.disaggregated.data.is.available, X2020) %>%
  rename(Deaths = Deaths.where.sex.disaggregated.data.is.available, GDP = X2020) %>%
  top_n(15)

GDP_plot <- ggplot(data = top_15_country_deaths, aes(x = Country,
                                    y = Deaths, color = GDP)) +
  geom_point() +
  labs(title = "Top 15 COVID Deaths by Country vs GDP", x = "Country", y = "Deaths from COVID") + theme(axis.text.x = element_text(angle = 90))

GDP_plot
``` 


```{r Relationships between Variables2, warning=FALSE, echo=FALSE} 
state_data <- raw_state_data %>% 
  group_by(location) %>% 
  filter(date == max(date)) %>% 
  select(location, total_boosters_per_hundred)

states <- states(cb = T)


states_merged <- geo_join(states, state_data, "NAME", "location")
states_merged <- subset(states_merged, !is.na(total_boosters_per_hundred))

pal <- colorNumeric("Blues", domain = states_merged$total_boosters_per_hundred)



map <- leaflet(states_merged) %>%
  addTiles() %>% 
  addPolygons(data = states_merged,
              color = "black",
              opacity = 1,
              fillColor = ~pal(states_merged$total_boosters_per_hundred),
              fillOpacity = 1, 
              weight = 1, 
              smoothFactor = 0.8,
              popup = ~total_boosters_per_hundred) %>% 
  addLegend("bottomleft", pal = pal, values = states_merged$total_boosters_per_hundred, 
                layerId = "colorLegend", title = "Booster Shot %", opacity = 1)

map

```


